Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-09-18 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

\begin{figure}[h]
  \label{fig:4_3}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_3.png}
  \caption{An «index» function}
\end{figure}

@@

Let us formulate the conditions of convergence for the algorithm in the form of the following theorem.
\begin{theorem}
Let the considered index algorithm be applied to the solving of problem \eqref{eq4:problem} and the following conditions be satisfied:
\begin{enumerate}
  \item each function $gj, 1\le j\le m+1$ satisfies Lipschitz condition with the constant $L_j$ in the interval $[a,b]$, i.e.,
  \[
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2 \in[a,b]
  \]
  \item the following inequalities hold for the quantities $\mu_\nu$ from \eqref{eq4:lip_est} starting from some step
  \[
  \r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
  \]
\end{enumerate}

@@

\section{Index algorithm taking into account the existence of -reserved solutions}

\section*{Index method with adaptive order of checking for constraints}
According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfillment of the problem constraints in the corresponding point of the search domain. The finding of the first violated constraint terminates the trial and initiates the transition to the next iteration. If a numeration of the constraints in the increasing order by the computational costs of their checking is introduced (i.e., if the simpler constraints are checked first) one can reduce essentially the overall computational costs in the course of optimization. The above is important for the problems, in which the estimating of the function value requires a considerable computational resources connected with a necessity of the numerical modelling of the behavior of the systems described by these functions (as it takes place, for example, in numerous problems of the optimal design).

@@

\section{A brief conclusion}
Constrained global optimization problems are under consideration. The method of penalty functions is one of the most popular numerical methods for solving the problems of this kind. Nevertheless, the method of penalty functions has a number of disadvantages. A new method for reduction of a constrained problem to an unconstrained one, so called \emph{index method}, has been considered in this chapter. This method, unlike the classical penalty function one, allows avoiding the problem of the parameter selection (like the penalty coefficients for the constraints). It is also allows solving the problems with partially defined objective function and constraints. This is quite natural for many applied problems, especially, in the optimal design of technical systems, because if some conditions of operation are not met, then some other characteristics of the system performance may not be defined.

>>>>

\begin{figure}[h]
  \label{fig:4_3}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_3.png}
  \caption{An ''index'' function}
\end{figure}

@@

Let us formulate the conditions of convergence for the algorithm in the form of the following theorem.
\begin{theorem}
Let the considered index algorithm be applied to the solving of problem \eqref{eq4:problem} and the following conditions be satisfied:
\begin{enumerate}
  \item each function $gj, 1\le j\le m+1$ satisfies Lipschitz condition with the constant $L_j$ in the interval $[a,b]$, i.e.,
  \[
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2 \in[a,b]
  \]
  \item the following inequalities hold for the quantities $\mu_\nu$ from \eqref{eq4:lip_est} starting from some step
  \[
  r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
  \]
\end{enumerate}

@@

\textbf{The results of the experiment.} Let us consider a problem from Example (\ref{ex4:problem}) as an illustration. Under assumption of partial computability the arcs of the problem functions corresponding to the domains $Q_j, 1\le j\le 3$, from \eqref{eq4:q} are presented in Fig. \ref{fig:4_6}.

Three methods have been used for solving this example:
\begin{enumerate}

  \item Scanning a uniform grid with the precision (distance between adjacent nodes) $\epsilon=10^{-5}$.
  \item Penalty function method, i.e., the minimization of the function
  \[
  p(x)=\varphi(x)+(\max\{g_1(x),0\})^2+(\max\{g_2(x),0\})^2
  \]
  using Algorithm of Global Search. The coordinates of the trial points executed by AGS in the process of problem solving are marked by the vertical strokes in Fig. \ref{fig:4_5}. The coordinates of the trials executed in the close points are marked by a dark rectangle.
  \item Index algorithm with the parameters $r_\nu=2, 1\le \nu \le 3$, and $\epsilon =10^{- 5}$. The trial point coordinates executed by the algorithm in the course of problem solving are marked by three rows of the vertical strokes in Fig. \ref{fig:4_6}. The strokes in the upper row correspond to the points with the index $\nu=1$, the ones of the second row --- to the points, the indices of which are equal to 2; and the points marked by the strokes in the bottom row are the feasible ones. The trial coordinates performed in the close points are marked by a dark rectangle.
\end{enumerate}

\begin{figure}[h]
  \label{fig:4_5}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_5.png}
  \caption{Minimization by means of Penalty Function Method}
\end{figure}

It is worth noting, that in the applied problems the estimates of the constraints as well as of the objective functions require considerable computation resources very often. The results presented in Table \ref{tab4:exp_results} confirm the effectiveness of the index scheme of the account of
constraints.

\begin{table}

\begin{center}
  \label{tab4:exp_results}
  \begin{tabular}{| l | c | c | c| }
    \hline
    & $k_1$ & $k_2$ & $k_3$ \\ \hline
    Scanning a uniform grid & 160000 & 90280 & 56476 \\ \hline
    Penalty Function Method &  375 & 375 & 375 \\ \hline
    Index method & 63 & 49 & 35 \\ \hline
  \end{tabular}
\end{center}
\end{table}

Here $k_i$ is the number of computations of the $i$ problem function, i.e., of the function $g_i(x )$.
\begin{figure}[h]
  \label{fig:4_6}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_6.png}
  \caption{The minimization of an ''index'' function}
\end{figure}

\section{Index algorithm taking into account the existence of $\epsilon$-reserved solutions}
Let us return to the consideration of a one-dimensional problem of the kind \eqref{eq4:problem}
\begin{displaymath}
  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
\end{displaymath}
In the case when the objective function and the left-side parts of the constraints are partially computable, it is necessary to use the form \eqref{eq4:problem2}.
\example
{
Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
}

\begin{figure}[h]
  \label{fig:4_7}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_7.jpg}
  \caption{The solution of a problem by Index Method}
\end{figure}

\begin{figure}[h]
  \label{fig:4_8}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_8.jpg}
  \caption{The problem solving by Index Method with the $\epsilon$-reservs}
\end{figure}

\section{Index method with adaptive order of checking for constraints}
According to the rules of the index method considered in the preceding subsections, every trial includes the successive check of the fulfillment of the problem constraints in the corresponding point of the search domain. The finding of the first violated constraint terminates the trial and initiates the transition to the next iteration. If a numeration of the constraints in the increasing order by the computational costs of their checking is introduced (i.e., if the simpler constraints are checked first) one can reduce essentially the overall computational costs in the course of optimization. The above is important for the problems, in which the estimating of the function value requires a considerable computational resources connected with a necessity of the numerical modelling of the behavior of the systems described by these functions (as it takes place, for example, in numerous problems of the optimal design).

@@

\section*{A brief conclusion}
Constrained global optimization problems are under consideration. The method of penalty functions is one of the most popular numerical methods for solving the problems of this kind. Nevertheless, the method of penalty functions has a number of disadvantages. A new method for reduction of a constrained problem to an unconstrained one, so called \emph{index method}, has been considered in this chapter. This method, unlike the classical penalty function one, allows avoiding the problem of the parameter selection (like the penalty coefficients for the constraints). It is also allows solving the problems with partially defined objective function and constraints. This is quite natural for many applied problems, especially, in the optimal design of technical systems, because if some conditions of operation are not met, then some other characteristics of the system performance may not be defined.
